Significant wind disturbances induced by giant dunes.

### 2021-11-26 Paper/main.tex
temp commit
First draft Introduction

\section{Introduction}
\section{Wind regimes across the Namib Sand Sea}

>>>>

\section{Introduction}

Whenever a flow encounters an obstacle, various types of interactions can arise depending the different time and lengths scales involved. In the case of atmospheric flows, this depends mainly on the part of the vertical structure of the atmosphere, schematically composed of a turbulent boundary layer topped by a turbulence-free part, with which the obstacle interacts.
%
At the largest scale, the feedback of mountains on the stratified flow of the free atmosphere results in wave generation as well as significant wind disturbances, such as foehn winds in the lee side \citep{refs}. Inside the boundary layer, the interaction between a turbulent flow and hilly surfaces is for example key to the understanding ocean surface wind-driven waves, or eolian beforms in desert \citep{Belcher1998, Sullivan2010, Courrech2015}.

Indeed, eolian sand dunes typically emerge from the interaction by the turbulent flow and the topography \citep{Charru2013}. Later on, when the dune reaches an intermediate size, it may induce large deflections that can significantly impact sediment pathways of coastal systems \citep{Hesp2015}. Likewise, the collective behavior of dune populations may also be affected by long-range interactions due to flow disturbances induced by each individual \citep{Smith2017, Bacik2020}. As the dunes increase in size by collisions and coarsening, they sometimes reach a giant size, comparable to the boundary layer depth, thus inducing interactions not only with the turbulent flow of the ABL, but also with the free atmosphere~\citep{andreotti2009}. However, the wind disturbances induced by these giant dunes have never been quantified.

The study of the impact of obstacles on the atmospheric flows allow its incorporation within meteorological numerical model. Therefore, they mainly become limited by the precision of the included topographical data, as well as the spatial grid of the model. For example, the latest climate reanalysis, ERA5-Land, is limited by its $9~\textup{km}$ spatial resolution, while including the data $30$-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007, munoz2021}. As such, it can not reproduce the flow disturbances induced by giant dunes, which have a typical length scale $\sim 1~\textup{km}$.

Here, we compare the wind predictions from the ERA5-Land dataset to local measurements in four different places across the Namib desert. In places with no significant topographies smaller than the model grid, we show that both wind datasets agree with each other. On the contrary, in places with giant dunes, we show that they may differ for some specific meteorological conditions, that we link to the circadian cycle of the ABL. We thus highlight the importance of the mid-scale topographies for local wind regimes, and its implications in the case of sand seas for smaller-scale eolian bedforms.

\section{Wind regimes across the Namib Sand Sea}
